### 2015-07-20 implementation-concretization.tex
Name change spack-paper-sc15 -> spack-sc15
Limitations section.
Concretization runtime section.

%!TEX root = spack-paper-sc15.tex

\subsection{Concretization}

@@

When the concretization process completes, it outputs a fully concrete spec DAG. 
Figure~\ref{fig:specs-mpileaks-concrete} shows a concrete DAG with architectures, 
compilers, versions, variants, and all dependencies resolved.
This fulfills Spack's guarantees.

\begin{figure}
	\centering
	\includegraphics[width=\columnwidth]{specs/mpileaks-concrete.pdf}
	\caption{
		Concretized spec from Figure~\ref{fig:specs-mpileaks}.
		\label{fig:specs-mpileaks-concrete}
	}
\end{figure}

@@

\subsubsection{Shared sub-DAGs}\label{sec:directory-layout}
\begin{figure}\centering
   \includegraphics[width=\linewidth]{specs/rpaths.pdf}
   \caption{
       {\tt mpileaks} built with {\tt mpich}, then {\tt openmpi}.
       \label{fig:reuse}
   }
\end{figure}

@@

\subsubsection{Reproducibility}
For reproducibility, and to preserve provenance, Spack stores a number of 
files in the installation directory that document how the installed package was
built.  These include the {\tt package.py} file used to build, a build log
containing output and error messages, and a file containing the complete
concrete spec for the package {\it and} its dependencies. The spec file can be
used later to reproduce the build, even if concretization preferences have changed.

>>>>

%!TEX root = spack-sc15.tex

\subsection{Concretization}
\label{sec:concretization}

@@

\begin{figure}
	\centering
	\includegraphics[width=\columnwidth]{specs/mpileaks-concrete.pdf}
	\caption{
		Concretized spec from Figure~\ref{fig:specs-mpileaks}.
		\label{fig:specs-mpileaks-concrete}
	}
\end{figure}
When the concretization process completes, it outputs a fully concrete spec DAG.
Figure~\ref{fig:specs-mpileaks-concrete} shows a concrete DAG with architectures,
compilers, versions, variants, and all dependencies resolved.
This fulfills Spack's guarantees.

@@

\subsubsection{Concretization Runtime}
\label{sec:concretization-overhead}

\begin{figure}
	\centering
	\includegraphics[width=\columnwidth]{figs/concretization-overhead/concretization-times.pdf}
	\caption{
		Concretization running time for 245 packages.
		\label{fig:concretization-time}
	}
\end{figure}

Figure~\ref{fig:concretization-time} shows the running time of concretization vs.
the size of a package DAG (in nodes).  We generated this plot by running {\tt concretize}
on all 245 packages in our internal version of Spack.  We see that for all but the 9 largest
packages, the algorithm takes less than 2 seconds on a Sandy Bridge cluster node, and less
than 1 second on a Haswell MacBook Pro.  For larger DAGs, we begin to see superlinear
scaling, but even for 50 nodes or more, concretization takes less than 6 seconds.
%
This performance is expected from a greedy, fixed-point method.
The algorithm repeats, evaluating conditions and adding dependencies and constraints, until
there are no more changes to be made.
%
%  Additionally, our current compiler concretization
% policy requires a search in the DAG; it prefers the compiler of the nearest ancestor
% if the current node's compiler is not set. 
%
More generally, concretization is an instance of the constraint
satisfaction problem, which is NP-complete. Our greedy solver is sufficient for the
packages we have built so far, and its running time is currently negligible. We do
not expect to see software packages with thousands of dependencies in the near
future, and we do not expect concretization to become a bottleneck, even if we
use a full constraint solver.

\subsubsection{Shared sub-DAGs}
\label{sec:directory-layout}

\begin{figure}\centering
   \includegraphics[width=\linewidth]{specs/rpaths.pdf}
   \caption{
       {\tt mpileaks} built with {\tt mpich}, then {\tt openmpi}.
       \label{fig:reuse}
   }
\end{figure}

@@

\subsubsection{Reproducibility}

For reproducibility, and to preserve provenance, Spack stores a number of
files in the installation directory that document how the installed package was
built.  These include the {\tt package.py} file used to build, a build log
containing output and error messages, and a file containing the complete
concrete spec for the package {\it and} its dependencies. The spec file can be
used later to reproduce the build, even if concretization preferences have changed.
